Are Greener Spaces Less Safe Places?

### 2024-04-19 other/datasheet/datasheet_template.qmd
Commit

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - TBD
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - TBD
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

>>>>

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze the relationship between the amount of green spaces in Toronto's neighborhoods and the corresponding crime rates. The goal was to understand if there is a correlation between green spaces and crime rates, which could help urban planners and policymakers in enhancing urban green spaces to potentially reduce crime rates.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Jena Shah created the dataset. She is a student at the University of Toronto.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - It was free.

### 2024-04-19 paper/paper.qmd
Commit

---
title: "Are Greener Spaces Less Safe Places?"
subtitle: "Exploring the Surprising Relationship Between Urban Green Spaces and Crime Rates in Toronto"
author: 
  - Jena Shah
thanks: "Code and data are available at: https://github.com/shahjena/greenspace_safeplace.git."
toc: true
date: today
date-format: long
abstract: "This study examines the relationship between green spaces and crime rates in Toronto's neighborhoods using gamma regression models. The analysis reveals a positive association between the amount of green spaces and crime rates, indicating that neighborhoods with more green spaces tend to have higher crime rates. This finding emphasizes the relation between urban green spaces and public safety, suggesting that the presence of green spaces may present challenges for crime prevention and urban planning. Understanding these dynamics can help the government and urban planners in their efforts to create safer and more sustainable cities."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

1.  HOOD_ID (Neighbourhood ID): This variable serves as a unique
    identifier for each neighborhood in Toronto, allowing us to
    differentiate between different areas in our analysis.

2.  AREA_NAME (Neighbourhood Name): This variable provides the name of
    each neighborhood, providing context and a more intuitive
    understanding of the data.

3.  Green.Spaces (Amount of Green Spaces): This variable quantifies the
    amount of green spaces in each neighborhood, measured in square
    kilometers. Green spaces include parks, forests, and other areas
    covered by vegetation.

4.  TOTAL_CRIMES (Total Crimes per Neighbourhood per 100000 People):
    This variable represents the total number of reported crimes per
    neighborhood, standardized to a rate per 100000 people. It provides
    a measure of the overall crime burden in each neighborhood,
    accounting for differences in population size.

These variables were selected based on their relevance to the study's
objective of examining the relationship between green spaces and crime
rates. The inclusion of the neighborhood ID and name allows us to
identify and differentiate between neighborhoods, while the Green.Spaces
variable provides a measure of the presence and extent of green spaces
in each area. The TOTAL_CRIMES variable provides a measure of the
overall crime rates, allowing us to assess the impact of green spaces on
crime in different neighborhoods. @fig-one shows the different variables
and we can see the right-skew in the crimes total, which can help us
decide what regression model to use in the next step.

@@

This study was done to investigate the relationship between the abundance of green
spaces and crime rates in urban neighborhoods in Toronto by using a
gamma regression model. This approach is suitable for analyzing count
data with a skewed distribution, such as crime rates, where the response
variable is non-negative and continuous.

@@

-   Dependent Variable: TOTAL_CRIMES (total crimes per neighborhood per
    100000 people).

@@

Overall, the gamma regression model is well-suited for analyzing the
relationship between green spaces and crime rates in Toronto's
neighborhoods, providing a statistically interpretable
approach to examining this important issue.

@@

In the results, I found that green spaces have a statistically significant relationship with crime rates in Toronto's neighborhoods. The gamma regression model showed that as the amount of green spaces in a neighborhood increases, the total crimes per 100000 people also tend to increase. Specifically, for every unit increase in green spaces (measured in square kilometers), there was a corresponding increase in total crimes.

@@

@fig-one illustrates this relationship with a scatterplot. It shows a concentration of data points at low to medium crime rates and low levels of green spaces. As the amount of green spaces increases, there are much fewer data points and they are more scattered, indicating higher crime rates in neighborhoods with more green spaces. The line of best fit in the scatterplot shows a nearly straight line, yet still highlighting the positive relationship between green spaces and crime rates.

@@

# Discussion
In this study, I analyzed the relationship between green spaces and crime rates in Toronto's neighborhoods using gamma regression models. The results suggest a positive relationship between the amount of green spaces and crime rates, indicating that areas with more green spaces tend to have higher crime rates. However, it is important to note that this relationship does not imply causation.

One of the key insights gained from this study is the potential impact of green spaces on crime rates in urban areas. The findings suggest that neighborhoods with more green spaces may face unique challenges in terms of crime prevention and urban planning. This knowledge can inform policymakers and urban planners in their efforts to create safer and more sustainable cities.

@@

Several limitations in this study should be acknowledged. Firstly, the green space data used in this analysis is from 2011, which may not accurately reflect the current green space distribution in Toronto. Additionally, the process of reducing the number of neighborhoods from 158 to 138 due to the removal the NA and 0 values could introduce bias into the analysis (so a total of 20 nighbourhoods were not considered in this study).

>>>>

---
title: "Are Greener Spaces Less Safe Places?"
subtitle: "Exploring the Surprising Relationship Between Urban Green Spaces and Crime Rates in Toronto"
author: 
  - Jena Shah
thanks: "Code and data are available at: https://github.com/shahjena/greenspace_safeplace.git."
toc: true
date: today
date-format: long
abstract: "This study examines the relationship between green spaces and crime rates in Toronto's neighbourhoods using gamma regression models. The analysis reveals a positive association between the amount of green spaces and crime rates, indicating that neighbourhoods with more green spaces tend to have higher crime rates. This finding emphasizes the relation between urban green spaces and public safety, suggesting that the presence of green spaces may present challenges for crime prevention and urban planning. Understanding these dynamics can help the government and urban planners in their efforts to create safer and more sustainable cities."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

1.  HOOD_ID (Neighbourhood ID): This variable serves as a unique
    identifier for each neighbourhood in Toronto, allowing us to
    differentiate between different areas in our analysis.

2.  AREA_NAME (Neighbourhood Name): This variable provides the name of
    each neighbourhood, providing context and a more intuitive
    understanding of the data.

3.  Green.Spaces (Amount of Green Spaces): This variable quantifies the
    amount of green spaces in each neighbourhood, measured in square
    kilometers. Green spaces include parks, forests, and other areas
    covered by vegetation.

4.  TOTAL_CRIMES (Total Crimes per Neighbourhood per 100000 People):
    This variable represents the total number of reported crimes per
    neighbourhood, standardized to a rate per 100000 people. It provides
    a measure of the overall crime burden in each neighbourhood,
    accounting for differences in population size.

These variables were selected based on their relevance to the study's
objective of examining the relationship between green spaces and crime
rates. The inclusion of the neighbourhood ID and name allows us to
identify and differentiate between neighbourhoods, while the Green.Spaces
variable provides a measure of the presence and extent of green spaces
in each area. The TOTAL_CRIMES variable provides a measure of the
overall crime rates, allowing us to assess the impact of green spaces on
crime in different neighbourhoods. @fig-one shows the different variables
and we can see the right-skew in the crimes total, which can help us
decide what regression model to use in the next step.

@@

This study was done to investigate the relationship between the abundance of green
spaces and crime rates in urban neighbourhoods in Toronto by using a
gamma regression model. This approach is suitable for analyzing count
data with a skewed distribution, such as crime rates, where the response
variable is non-negative and continuous.

@@

-   Dependent Variable: TOTAL_CRIMES (total crimes per neighbourhood per
    100000 people).

@@

Overall, the gamma regression model is well-suited for analyzing the
relationship between green spaces and crime rates in Toronto's
neighbourhoods, providing a statistically interpretable
approach to examining this important issue.

@@

In the results, I found that green spaces have a statistically significant relationship with crime rates in Toronto's neighbourhoods. The gamma regression model showed that as the amount of green spaces in a neighbourhood increases, the total crimes per 100000 people also tend to increase. Specifically, for every unit increase in green spaces (measured in square kilometers), there was a corresponding increase in total crimes.

@@

@fig-one illustrates this relationship with a scatterplot. It shows a concentration of data points at low to medium crime rates and low levels of green spaces. As the amount of green spaces increases, there are much fewer data points and they are more scattered, indicating higher crime rates in neighbourhoods with more green spaces. The line of best fit in the scatterplot shows a nearly straight line, yet still highlighting the positive relationship between green spaces and crime rates.

@@

# Discussion
In this study, I analyzed the relationship between green spaces and crime rates in Toronto's neighbourhoods using gamma regression models. The results suggest a positive relationship between the amount of green spaces and crime rates, indicating that areas with more green spaces tend to have higher crime rates. However, it is important to note that this relationship does not imply causation.

One of the key insights gained from this study is the potential impact of green spaces on crime rates in urban areas. The findings suggest that neighbourhoods with more green spaces may face unique challenges in terms of crime prevention and urban planning. This knowledge can inform policymakers and urban planners in their efforts to create safer and more sustainable cities.

@@

Several limitations in this study should be acknowledged. Firstly, the green space data used in this analysis is from 2011, which may not accurately reflect the current green space distribution in Toronto. Additionally, the process of reducing the number of neighbourhoods from 158 to 138 due to the removal the NA and 0 values could introduce bias into the analysis (so a total of 20 nighbourhoods were not considered in this study).
